Intentional release of native species undermines ecological stability

### 2022-10-20 rmd/manuscript.Rmd
Update manuscript.Rmd

suppressMessages(source(here::here("code/set_rmd.R")))
df_sim <- list_sim[[2]]

@@

When carrying capacity was small ($K=$ `r min(list_param[[2]]$k)`), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure** \@ref(fig:figure-layout)**A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure** \@ref(fig:figure-layout)**A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals, as reported in a previous theoretical study [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure** \@ref(fig:figure-layout)**A**). Combined, the total community density decreased more sharply than individual species groups (**Figure** \@ref(fig:figure-layout)**A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure** \@ref(fig:figure-layout)**A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S3-S6**).

@@

The intentional release had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7-S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. The weak-competition scenario is likely in re-introduction programs of extirpated species or population enhancement of long-lived endangered species. In fact, some of the best evidence for successful intentional release comes from re-introduction efforts [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the once-extirpated Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial re-introduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. However, the preponderance of unsuccessful re-introduction programs implies that the theoretical requirements for successful intentional release are not always met [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into unsuitable habitats with compromised carrying capacities [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that such programs could rather impair biodiversity without noticeable demographic benefit to the enhanced species. Pre-release examination and restoration of environmental capacity may be key to successful release programs with minimal impacts on other community members.

@@

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with the intensive release (**Figure** \@ref(fig:figure-layout)**B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure** \@ref(fig:figure-coef); see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures** \@ref(fig:figure-layout)**B and** \@ref(fig:figure-coef)), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures** \@ref(fig:figure-layout)**B and** \@ref(fig:figure-coef)**;** see **Tables S6-S8** for full statistics). In the meantime, the SDs had vague relationships with the intentional release (**Figures** \@ref(fig:figure-layout)**B and** \@ref(fig:figure-coef)).

@@

It is noteworthy that the scale of the release program at the protected watersheds (annual average maximum = `r op(max(df_stock_mu$mean_stock), 2)` million fish) is fairly small when compared with those for other fishery resources [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]. For example, 649.1 million fish are released annually to supplement the pink salmon stock at the Prince William Sound in Alaska, where wild populations seem to be severely compromised due to competition with hatchery fish [@amorosoMeasuringNetBiological2017]. Similarly, large-scale releases of Black sea bream (on average, 0.64 million juveniles annually) has been conducted in Kagoshima Bay (Japan) since 1974 with associated declines of wild stocks and genetic diversity [@kitadaRigorousMonitoringLargescale2019]. Although we do not know the carrying capacities of the recipient systems, it is clear that our results should not be viewed as an exceptional case. In support of this view, a meta-analysis revealed that only a few massive release programs in fisheries led to an increased abundance of the focal species [@arakiHatcheryStockingHelp2010]. Yet, none of those studies statistically quantified the community-level impact. Future studies addressing this knowledge gap are therefore desired to generalize our findings.

@@

```{r figure-alpha, fig.cap="Estimated competition coefficients of the multi-species Ricker model. Dots represent a given pair of taxa, and box plots show median (vertical center line) and quartiles (box limits) with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range. The density plot on the top shows the overall distribution of the competition coefficient. Colors distinguish taxa."}

>>>>

suppressMessages(source(here::here("code/set_rmd.R")))

@@

When carrying capacity was small ($K=$ `r min(list_param[[2]]$k)`), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure** \@ref(fig:figure-layout)**A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure** \@ref(fig:figure-layout)**A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals, as reported in a previous theoretical study [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure** \@ref(fig:figure-layout)**A**). Combined, the total community density decreased more sharply than individual species groups (**Figure** \@ref(fig:figure-layout)**A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure** \@ref(fig:figure-layout)**A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S3 -- S6**).

@@

The intentional release had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7 -- S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. The weak-competition scenario is likely in re-introduction programs of extirpated species or population enhancement of long-lived endangered species. In fact, some of the best evidence for successful intentional release comes from re-introduction efforts [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the once-extirpated Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial re-introduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. However, the preponderance of unsuccessful re-introduction programs implies that the theoretical requirements for successful intentional release are not always met [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into unsuitable habitats with compromised carrying capacities [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that such programs could rather impair biodiversity without noticeable demographic benefit to the enhanced species. Pre-release examination and restoration of environmental capacity may be key to successful release programs with minimal impacts on other community members.

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with the intensive release (**Figure** \@ref(fig:figure-layout)**B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure** \@ref(fig:figure-coef); see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures** \@ref(fig:figure-layout)**B and** \@ref(fig:figure-coef)), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures** \@ref(fig:figure-layout)**B and** \@ref(fig:figure-coef)**;** see **Tables S6 -- S8** for full statistics). In the meantime, the SDs had vague relationships with the intentional release (**Figures** \@ref(fig:figure-layout)**B and** \@ref(fig:figure-coef)).

@@

@@

It is noteworthy that the scale of the release program at the protected watersheds (annual average maximum = `r op(max(df_stock_mu$mean_stock), 2)` million fish) is fairly small when compared with those for other fishery resources [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]. For example, 649.1 million fish are released annually to supplement the pink salmon stock at the Prince William Sound in Alaska, where wild populations seem to be severely compromised due to competition with hatchery fish [@amorosoMeasuringNetBiological2017]. Similarly, the large-scale release of Black sea bream has been conducted in Kagoshima Bay (Japan) since 1974 (the yearly-average of 0.64 million juveniles) with associated declines of wild stocks and genetic diversity [@kitadaRigorousMonitoringLargescale2019]. Although we do not know the carrying capacities of the recipient systems, it is clear that our results should not be viewed as an exceptional case. In support of this view, a meta-analysis revealed that only a few massive release programs in fisheries led to an increased abundance of the focal species [@arakiHatcheryStockingHelp2010]. Yet, none of those studies statistically quantified the community-level impact. Future studies addressing this knowledge gap are therefore desired to generalize our findings.

@@

```{r figure-alpha, fig.cap="Estimated competition coefficients of the multi-species Ricker model. Dots represent a given pair of taxa, and box plots show median (vertical center line) and quartiles (box limits) with whiskers extending up to the most extreme data points that are within ±1.5 interquartile range. The density plot on the top shows the overall distribution of the competition coefficient. Colors distinguish taxa. The vertical broken line denotes the overall median value."}
